### 2017-02-02 2017mergingmerges/section/01-Introduction.tex
nothing useful

Suppose we have two large arrays containing textual identifiers, and wish to perform some analysis on them. Both arrays are sorted, but may contain duplicate identifiers. 
We wish to retrieve the unique set of identifiers from the first array, as well as produce the union of both ararys. Can we perform both of these tasks at once, without reading the input arrays multiple times?
\begin{code}
  uniquesUnion : Array Text -> Array Text -> (Array Text, Array Text)
  uniquesUnion arr1 arr2
   = let  uniques1 = group arr1
          merged   = merge arr1 arr2
          union    = group merged
     in   (uniques1, union)
\end{code}

@@

The above example is seeming innocuous, but existing fusion transformations cannot handle it. Stream fusion~\cite{coutts2007stream} and other forms of shortcut fusion~\cite{jones2001playing} cannot avoid performing duplicate memory access when a particular array or intermediate result is used by multiple consumers. We refer to this situation as a \emph{split} --- when the example above is viewed as a data-flow network the flow from the argument array @arr1@ is split into both the @group@ and @merge@ consumers. Synchronous dataflow languages~\cite{mandel2010lucy} and data flow inspired array fusion~\cite{lippmeier2013data} also cannot fuse this example because they do not support combinators with multiple inputs that have value-dependent access patterns. We refer to such as situation as a \emph{join} --- in the example above the @merge@ combinator expresses a join in the data-flow graph. Polarized data flow fusion~\cite{lippmeier2016polarized} \emph{is} able to fuse this case, but requires the program to be rewritten to use explicitly polarized stream types, rather than the more familiar arrays.

>>>>

Suppose we have two large arrays containing textual identifiers, and wish to perform some analysis on them. Both arrays are sorted, but may contain duplicate identifiers. 
We wish to retrieve the unique set of identifiers from the first array, as well as produce the union of both arrays. Can we perform both of these tasks at once, without reading the input arrays multiple times?
\begin{code}
  uniquesUnion : Array Text -> Array Text -> (Array Text, Array Text)
  uniquesUnion arr1 arr2
   = let  uniques1 = group arr1
          merged   = merge arr1 arr2
          union    = group merged
     in   (uniques1, union)
\end{code}

@@

The above example seems innocuous, but existing fusion transformations cannot handle it.
Stream fusion~\cite{coutts2007stream} and other forms of shortcut fusion~\cite{jones2001playing} cannot avoid performing duplicate memory access when a particular array or intermediate result is used by multiple consumers.
We refer to this situation as a \emph{split} --- when the example above is viewed as a data-flow network the flow from the argument array @arr1@ is split into both the @group@ and @merge@ consumers.
Recent work on stream fusion by \citet{kiselyov2016stream} uses staged computation to ensure all combinators are inlined, but in the presence of splits this causes excessive inlining which duplicates work.
Synchronous dataflow languages~\cite{mandel2010lucy} and data flow inspired array fusion~\cite{lippmeier2013data} also cannot fuse this example because they do not support combinators with multiple inputs that have value-dependent access patterns.
We refer to such as situation as a \emph{join} --- in the example above the @merge@ combinator expresses a join in the data-flow graph.
Polarized data flow fusion~\cite{lippmeier2016polarized} \emph{is} able to fuse this case, but requires the program to be rewritten to use explicitly polarized stream types, rather than the more familiar arrays.

\amos{The split/join distinction is actually more about pushes vs pulls, as \emph{some} joins can be expressed with synchronous dataflow eg non-deterministic merge}
